An agent-based model of military mechanization

### 2018-05-07 paper2/paper2.tex
foo

\section{Introduction}

@@

Importantly, if the findings of \cite{sechser2010army} are correct and states
largely determine their own mechanization levels on the basis of what other
states are doing, a positive feedback loop exists. State A has high
mechanization, causing enemy State B to raise their mechanization, causing State
A to raise their mechanization... and so on. This feedback loop---often called
the ``security dilemma''---has been extensively studied, including by Thomas Schelling, one of the pioneers
of agent-based modeling (\cite{schelling1960strategy};
\cite{schelling2006micromotives}).

@@

There is one agent type: states. Agents exist at a fixed location and do not
move. The composition of agents changes throughout the

@@

Agents exist in a spatial environment. Spatiality is represented by neighbor
relationships rather than a raster or vector GIS map. Two states are defined as
being neighbors if they share a contiguous land border or a stretch of water
less than 400 miles. Every agent has an attribute \texttt{neighbor\_list}, a
list in which it stores the names of neighboring agents for that time-step.

\subsection{Initialization and Time}

\subsection{Dynamics and Interactions}

>>>>

\begin{centering}
	Word count: XXXX
\end{centering}
\section{Introduction}

@@

Importantly, the findings of \cite{sechser2010army} imply a positive feedback
loop. State A has high
mechanization, causing enemy State B to raise their mechanization, causing State
A to raise their mechanization... and so on. Such feedback loops are a hallmark
of complex systems and again are well-suited for representation by agent-based
modeling. This specific one---states changing their military in response to
other states change in military in response to... and so on---feedback loop---is often called
the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
of agent-based modeling (\cite{schelling1960strategy};
\cite{schelling2006micromotives}).

@@

There is one agent type: states. Agents exist at a fixed location and do not
move. The set of agents encompasses all diplomatically-recognized 
states in the world at a particular point in time, minus microstates (defined as
those with a population less than 750,000).

Agents have attributes. \textit{Name} is self-explanatory.
\textit{Mechanization} is the agent's level of mechanization at a given
timestep. Mechanization is calculated as the number of armored vehicles per 100
soldiers, per \cite{sechser2010army}. \textit{Neighbors} is a list containing
the \textit{Names} of an agent's geographic neighbors at a given point in
time---more on this in the subsequent \textbf{Environment} section. \textit{NeighborsMech} is the
average mechanization level of an agent's \textit{Neighbors} at a given point in
time. \textit{Enemies} is a list containing the \textit{Names} of an agent's
enemies at a given point in time. An enemy is an agent with which the agent in
question has had a militarized interstate dispute in the preceding ten years,
per the Correlates of War (COW) Militarized Interstate Disputes Database (MIDB)
(\cite{cow_midb}). \textit{EnemiesMech} is the average mechanization score of
enemies. Agents also have attributes that affect their
perception of other agents and cognition for changing their own behavior. These
attributes are discussed in the \textbf{Cognition and Actions} section.

@@

Agents exist in a spatial environment. Spatiality is represented by neighbor
relationships. Two states are defined as
being neighbors if they share a contiguous land border or a stretch of water
less than 400 miles. Data on neighbor relationships is taken from the Correlates
of War (COW) Direct Continuity dataset (\cite{cow_contiguity}).
Modeling spatiality as dyadic neighbor relationships, rather than a more
detailed raster or vector GIS implementation, is justified on the basis that the
academic literature on mechanization suggests that two states neighboring each
other is the important factor, not more detailed geographic information (e.g.,
topographical features along the border). As such, this paper's modeling choice
captures the phenomenon of interest while avoiding the overhead of a full-blown GIS implementation.

\subsection{Cognition}

Agents have global perception of other agents: they can ``see'' the
mechanization state of all other agents. This modeling choice is reflective of
the real world, in which states devote significant resources to stay informed of
the military posture of others. Agent cognition centers on 
updating their own mechanization in response to the external security
environment. This occurs in three ways. One, if an agent's
\textit{NeighborsMech} increases---i.e., the agent's neighbors become more
mechanized---the agent will increase its own mechanization by some proportional
amount controlled by the agent's unique \textit{NeighborMechSensitivity}
attribute, which is heterogenous across agents (more on this below).
Two, if an agent's
\textit{EnemiesMech} increases---i.e., the agent's enemies become more
mechanized---the agent will increase its own mechanization by some proportional
amount controlled by the agent's unique \textit{EnemyMechSensitivity} attribute,
again heterogenous across agents. Three,
agents occasionally suffer an ``anti-mechanization lesson'': defeat in a
counter-insurgency conflict. Such events are scripted to occur for agents and
years as per the historical record: e.g., the Soviet Union experienced such a
lesson in the 1980s in Afghanistan. An agent experiencing such an event will
decrease their own mechanization by some amount controlled by the agent's
unique \textit{LessonSensitivity} attribute, again heterogenous across agents.

An important note is that these three factors---neighbors, enemies, and
defeat in counter-insurgency---correspond to the factors identified by 
\cite{sechser2010army} as most important for mechanization. Because the research
goal is essentially to see whether their findings stand up to agentization,
it is critical to make agent cognition work in about the way that their findings
predict. Inclusion of other factors not identified as important for
mechanization---e.g., alliances, economic factors,
culture, etc.---would be interesting but counter-productive to the research
question. They are discussed along with this model's findings.

Data for \textit{NeighborsMechSensitivity}, \textit{EnemiesMechSensitivity}, and
\textit{LessonSensitivity} come from \cite{sechser2010army}. Specifically, they
estimate coefficients and standard errors for these three factors, thus defining
a statistical distribution for each. Upon instantiation, every agent makes a
random draw from each of the three distributions, the resulting values of are
taken for these three attributes. This introduces a stochastic element to the
model and variation across runs, providing a richer body of data to draw upon.

\subsection{Time}

The model starts in 1979 and runs in 2-year time increments until 2001. This
timespan and periodicity reflects data availability: a longer time period and
more frequent updates would be ideal but is not practically achievable. Agent
attributes are initialized, and then then the model iterates through timesteps.

In discussing what happens each timestep, it is useful to distinguish between
things that depend on agents' actions and those that do not.

Some changes are deterministically programmed into the model and do not depend on
agent behavior. Every timestep, neighbor relationships and enemy relationships
are updated in accordance with the historical record. The occurrence of
counter-insurgency defeats also is hardwired into the model and does not depend
on agent behavior within the model.

Other changes do depend on agent behavior. Agent activation order in a
particular timestep is random. Every timestep, every agent perceives the world
aorund them; recalculates \textit{NeighborsMech} and \textit{EnemiesMech};
updates their mechanization level according to \textit{NeighborMechSensitivity}
and \textit{EnemyMechSensitivity}; and updates their mechanization level
according to \textit{LessonSensitivity} if an anti-mechanization lesson occured.

\subsection{Output Data}
